---
title: "Stevia Microbiota"
author: "Edna Chiang"
date: "October 30, 2018"
output:
  html_document:
    toc: true
    toc_float: true
---


```{r, include=F}
# Load Libraries
library(ape)
library(dplyr)
library(ggplot2)
library(gplots)
library(phangorn)
library(tidyr)
library(vegan)
library(phyloseq)
library(grid)
library(pgirmess)
library(lme4)
library(kableExtra)
theme_set(theme_bw())
set.seed(1)
```

# Take-Away Points  

* There is little difference between High Fat, Saccharin, and Stevia samples.
* The majority of differences are between Low Fat and the other 3 treatments (all of which include a high fat diet)
  
  
# Description  
This is a quick summary of the follow-up analyses to Sarah's thesis that I've performed.  
A few differences between Sarah's analysis and my analysis:  

* I perform my analysis at the OTU-level (97% sequence similarity, akin to species-level)
  

* I included only individuals with both pre-treatment and post-treatment samples because I run a few paired tests and want to keep the samples consistent throughout all analyses. I removed 4 individuals, resulting in the following final sample count:  
```{r, echo=F}
samps <- data.frame(rep("NA",4))
colnames(samps) <- "Treatment"
samps$Treatment <- c("High Fat", "Low Fat", "Saccharin", "Stevia")
samps$N <- c(9, 10, 9, 8)
samps$Female <- c(4, 5, 5, 4)
samps$Male <- c(5, 5, 4, 4)
kable(samps, "html") %>% kable_styling(full_width=F)
```
  
  
# Phylum-Level Composition  
The phyla are listed in order of highest overall relative abundance to lowest overall relative abundance. Phyla with a total relative abundance of < 0.01% were removed. Error bars represent the interquartile range.  

![](plots/phyla.tiff)  
  
Dominant phyla include Firmicutes and Bacteroidetes, with lesser contributions by Verrucomicrobia and Tenericutes, and minor contributions by Proteobacteria, Actinobacteria, unclassified bacteria, and Cyanobacteria.  
  
# Within-Sample Diversity (Alpha)  
  
![](plots/adiv.tiff)  

 
Chao is an estimate of richness (# unique OTUs exist in a particular sample).  

* Post-treatment groups are significantly different (adj p-val = 0.004). Specifically:
    + High Fat vs. Low Fat
    + Low Fat vs. Saccharin
* High Fat Pre-treatment and Post-treatment are significantly different (adj p-val = 0.018)

Shannon's Index is an estimate of diversity, taking into account both the presence/absence and abundance of OTUs.   

* No significant differences.  
* This suggests that differences in within-sample diversity are driven by presence/absence of OTUs rather than their relative abundances.  
  
# Between-Sample Diversity (Beta)  
I used the Bray-Curtis dissimilarity metric, which considers both OTU presence/absence and weight (abundance).  
  
In the ordination, each point is 1 sample. The distance betwen points is inversely related to how similar the samples are.  
Closer points = more similar.  
Farther points = less similar.  

![](plots/nmds.tiff)

  
Pre-treatment vs. Post-treatment samples visually cluster distinctly for all samples except Low Fat.  
Within each treatment, Pre vs. Post are significantly different (all adj pvals = 0.005).  
  
To determine to what extent the categorical factors contribute to between-sample differences, I used PERMANOVA to evaluate:  

* Time
    + Pre-treatment vs. Post-treatment
* Treatment
    + High Fat, Low Fat, Saccharin, Stevia
* Sex
    + Female, Male
* Individual
    + My attempt to account for the paired nature of the data

```{r, echo=F}
samps <- data.frame(rep("NA",5))
colnames(samps) <- "Factor"
samps$Factor <- c("Time", "Treatment", "Sex", "Individual", "Residual")
samps$pval <- c(0.001, 0.001, 0.001, 0.907, NA)
samps$R2 <- c(0.258, 0.123, 0.074, 0.230, 0.349)

kable(samps, "html") %>% kable_styling(full_width=F)
```
  

* Time accounts for the most variation in the data, followed by Treatment and Sex
* Individual accounted for quite a bit of the data variation, but was not significant (likely because we have a small n (2) for each individual)
  
Pre-treatment:  

* There is no significance between treatments.  
* There *is* a significant difference between sex (adj pval = 0.005).


Post-treatment:  

* Unlike in Pre-treatment samples, sex is not significant in Post-treatment samples (adj p-val = 0.326)
* Low Fat is significantly different from all other treatments (all adj pvals = 0.005)
    + This is true overall, and within sex
* Saccharin vs. Stevia trend towards a difference (adj pval = 0.063)
    + This trend is seen within Females (adj pval = 0.063), but not within Males (adj pval = 0.571)

  
  
# Important OTUs  

I identified the OTUs that contribute the most to variation in between-sample diversity using pairwise comparisons. These OTUs are the most abundant and/or variable in the samples.  
This test (similarity percentages (SIMPER)) is run with pairwise comparisons.  
  
![](plots/hm.tiff)
  

## Male vs. Female   
These results are for Pre-treatment only.  
No OTUs were detected within Post-treatment samples overall.  
```{r, echo=F}
tab <- data.frame(rep("NA", 5))
tab[,1] <- c(9.74, 7.97, 7.06, 5.85, 5.58)
tab$Adj_Pval <- c(5.31e-06, 3.77e-05, 0.0075, 0.0002, 0.0001)
tab$Phylum <- c("Firmicutes", "Bacteroidetes", "Firmicutes", "Firmicutes", "Firmicutes")
tab$Class <- c("Clostridia", "Bacteroidia", "Clostridia", "Erysipelotrichia", "Clostridia")
tab$Order <- c("Clostridiales", "Bactteroidales", "Clostridiales", "Erysipelotrichales", "Clostridiales")
tab$Family <- c("Ruminococcaceae", "S24-7", "Lachnospiraceae", "Erysipelotrichaceae", "Lachnospiraceae")
tab$Genus <- c("Ruminiclostridium", "Unclassified", "Unclassified", "Unclassified", "Tyzzerella")
tab$Species <- c("OTU 6", "OTU 13", "OTU 5", "OTU 11", "OTU 8")
tab$Female_RelAbund <- c(12.34, 0.17, 3.86, 8.14, 9.18)
tab$Male_RelAbund <- c(0.01, 10.22, 11.42, 1.44, 2.29)
colnames(tab) <- c("Variation Explained (%)", "Adj Pval", "Phylum", "Class", "Order", "Family", "Genus", "OTU", "Female Rel Abund (%)", "Male Rel Abund (%)")
kable(tab, "html") %>% kable_styling(full_width=F)
```